### 2015-07-06 conclusion.tex
Word => TeX formatting
Authors, fixed-width fonts, citations, dumb quotes.

In this paper, we have described two effective algorithms for extracting date format strings from a small set of samples, one using a minimum descriptive length approach and one using natural language techniques. Both algorithms are accurate enough to be used automatically without user involvement. The MDL algorithm is also fast enough to deploy in an interactive environment.

While validating the algorithms on a large corpus, we also found that the number of distinct formats in the wild is surprisingly high, and demonstrates the wisdom of including general-purpose date parsing functions in data visualization tools, data cleaning tools and RDBMSes. In particular, it is interesting to note that the most prominent open source RDBMSes (MySQL and Postgres) both have a built-in version of \texttt{DATEPARSE}, possibly reflecting that this is a common need that gets implemented when users are empowered to add to a code base.