### 2015-07-06 futurework.tex
More formatting
Redo all the quotes (I’d forgotten how annoying TeX’s quote mark
conventions are). Format e.g. everywhere (next time I’ll define a
macro…)

While we are satisfied with the results so far, we found a number of date formats that we were not able to handle with these techniques. Many columns contained multiple formats, and we would like to recognize this situation and generate predicate calculations to increase the accuracy of the results and thereby make the results even more seamless for the user. Other columns contain date ranges, which we would like to handle by generating multiple calculations, possibly by combining regular expressions with date parsing.

In this paper, we have only considered the parsing of strings, but dates are often formatted as integers (\textit{e.g.} \texttt{201507016}). It is significantly faster to decompose integers into date parts using arithmetic operations (\textit{e.g.} mod and div) than by using slower, locale-sensitive string parsing functions. Moreover, the number of possible formats is low enough that we may be able to enumerate them. Timestamps are another form of numeric date for which we intend to automate the preparation.

In the course of our research, we have also identified a number of date part variants (\textit{e.g.} ordinal dates, four-letter month abbreviations, alternate meridian markers and postfix quarter syntax) that we would like to commend to the ICU project and maybe provide implementations for.