### 2015-07-14 relatedwork.tex

% vidya: this probably should be its own section
% hawkfish: Not sure it is big enough, but I can see the logic.

Data preparation has been considered an analytic bottleneck since at least the description of the Potter's Wheel system~\cite{Raman:2001}. Since then, several other interactive data preparation systems have been proposed, including Data Wrangler~\cite{Kandel:2011} and Google Refine~\cite{Refine}. While effective, these systems all make assumptions about possible date formats, which we suggest are too restrictive for real world data.

Various approaches have been described for deriving regular expressions, and a good overview is provided by Li et al. in their paper on the ReLIE system for deriving regular expressions given a starting expression provided by a domain expert~\cite{Li:2008}. The Minimum Descriptive Length technique first described in Rissanen~\cite{Rissanen:1978} was used in~\cite{Raman:2001} to generate regular expressions.

There are several bodies of research on developing semantic parsers and grammars for interpreting time and date expressions. Lee \textit{et al} use a combinatory categorical grammar by combining the TIMEX3 standard~\cite{timex3} with contextual cues such as document creation time to determine the reference for parsing time expressions such as `2nd Friday of July'~\cite{LeeADZ14}.

\vidya{Should this be preceded by a bulleted list of contributions, under the section Contributions?}
%\subsection{Overview}
The rest of this paper is organized as follows: The next section introduces the parameters of the problem space. The following two sections describe the two different algorithms, one using Minimum Descriptive Length and the other using Natural Language Processing. In section 5, we evaluate the algorithms on a corpus of 30K columns, both by sampling the outputs manually and then by using the algorithms to validate each other. We then discuss future work in section 6 and conclude in section 7.