### 2016-04-28 cikm2016/cikm_dateparse.tex
Some rewording in the abstract and removal of orphan words.

% THIS IS SIGPROC-SP.TEX - VERSION 3.1
% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
% APRIL 2009
%
% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
% LaTeX2e document class file for Conference Proceedings submissions.
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) Page numbering
% ---------------------------------------------------------------------------------------------------------------
% It is an example which *does* use the .bib file (from which the .bbl file
% is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission,
% you need to 'insert'  your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% Questions regarding SIGS should be sent to
% Adrienne Griscti ---> griscti@acm.org
%
% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
% Gerald Murray ---> murray@hq.acm.org
%
% For tracking purposes - this is V3.1SP - APRIL 2009

%\documentclass{acm_proc_article-sp}
%\documentclass[conference]{IEEEtran}
\documentclass{sig-alternate-05-2015}
%
\usepackage{url}
\usepackage{syntax}
\usepackage{amsmath}% http://ctan.org/pkg/amsmath
\usepackage{mathtools}
\usepackage{color}
\usepackage{caption} 
\usepackage{tikz}
\captionsetup[table]{skip=10pt}

%
\def\sharedaffiliation{%
\end{tabular}
\begin{tabular}{c}}
%
\def\eg{\textit{e.g.} }
\def\ie{\textit{i.e.} }
\def\dateparse{\mbox{\texttt{DATEPARSE}}}
\hyphenation{time-stamp}

% to add comments
\newcommand{\vidya}[1]{\textcolor{red}{(vidya: #1)}}
\newcommand{\richard}[1]{\textcolor{green}{(richard: #1)}}
\newcommand{\dan}[1]{\textcolor{blue}{(dan: #1)}}
\newcommand\functionname[1]{\mbox{\texttt{#1}}}

% formatting math symbols
\newcommand{\Mod}[1]{\ (\text{mod}\ #1)}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

% tightening the whitespace for itemize
\newenvironment{tight_itemize}{\begin{itemize} \itemsep
-2pt}{\end{itemize}}

%
\begin{document}

\title{Does Anybody Really Know What Time It Is?\\
Automating the Extraction of Date Scalars}
\input{anonymous}

\maketitle
\begin{abstract}
Interactive visual analytics is an effective approach for the timely analysis of big data. Users who are already engaged in interactive data analysis prefer to stay in the visualization environment for unavoidable data cleaning and preparation tasks in order to preserve their analytic flow. This has led visualization environments to include simple data preparation functions such as scalar parsing, pattern matching and categorical binning. One of the most common scalar parsing tasks is extracting date and time data from string representations. Several relational database systems include date parsing ``mini-languages'' to cover the wide range of possible formats, but analysis of user data from one visualization system shows that the parsing language syntax is difficult for users to master without disturbing their analytic flow.

In this paper, we present two algorithms for automatically deriving date formats from a column of data with minimal user disruption, one based on minimal entropy representations and another based on natural language processing techniques. Both have accuracies of over 95\% on a large corpus of date columns extracted from an online data repository and are in substantial agreement with each other. One approach is also fast enough to produce results within the user's perceptual threshold, making it suitable for interactive work.
\end{abstract}

%Looks like they don't use them.
%\vidya{Need to figure out the IEEE equivalent for categories and terms}
%D.3.3 [Data Processing]: Data Cleaning, Natural Language Processing.
% A category with the (minimum) three required fields
%\category{D.3.3}{Data Processing}{Data Cleaning, Natural Language Processing}
%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

%\terms{Algorithms, Performance}

\section{Introduction}
\input{intro}

\section{Background}
\input{background}

\section{Minimal Descriptive Length}
\input{mdl}

\section{Natural Language Processing}
\input{nlp}

\section{Experiments}
\input{testing}

\section{Related Work}
\input{relatedwork}

\section{Future Work}
\input{futurework}

\section{Conclusion}
\input{conclusion}

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
\input{credits}

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{dateparse}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ``.bib'' file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%

%\balancecolumns
% That's all folks!
\end{document}