### 2016-05-02 cikm2016/cikm_dateparse.tex
Add cross checking to abstract

% THIS IS SIGPROC-SP.TEX - VERSION 3.1
% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
% APRIL 2009
%
% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
% LaTeX2e document class file for Conference Proceedings submissions.
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) Page numbering
% ---------------------------------------------------------------------------------------------------------------
% It is an example which *does* use the .bib file (from which the .bbl file
% is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission,
% you need to 'insert'  your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% Questions regarding SIGS should be sent to
% Adrienne Griscti ---> griscti@acm.org
%
% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
% Gerald Murray ---> murray@hq.acm.org
%
% For tracking purposes - this is V3.1SP - APRIL 2009

%\documentclass{acm_proc_article-sp}
%\documentclass[conference]{IEEEtran}
\documentclass{sig-alternate-05-2015}
%
\usepackage{url}
\usepackage{syntax}
\usepackage{amsmath}% http://ctan.org/pkg/amsmath
\usepackage{mathtools}
\usepackage{color}
\usepackage{caption} 
\usepackage{tikz}
\usepackage{enumitem}
\captionsetup[table]{skip=10pt}

\newenvironment{tight_itemize}{\begin{itemize} \itemsep
-2pt}{\end{itemize}}
%
\def\sharedaffiliation{%
\end{tabular}
\begin{tabular}{c}}
%
\def\eg{\textit{e.g.\@} }
\def\ie{\textit{i.e.\@} }
\def\cf{\textit{cf.\@} }
\def\etal{\textit{et al.\@} }
\def\naive{na\"ive}
\def\Naive{Na\"ive}
\def\dateparse{\mbox{\texttt{DATEPARSE}}}
\hyphenation{time-stamp}

% to add comments
\newcommand{\vidya}[1]{\textcolor{red}{(vidya: #1)}}
\newcommand{\richard}[1]{\textcolor{green}{(richard: #1)}}
\newcommand{\dan}[1]{\textcolor{blue}{(dan: #1)}}
\newcommand\functionname[1]{\mbox{\texttt{#1}}}

% formatting math symbols
\newcommand{\Mod}[1]{\ (\text{mod}\ #1)}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

% Alter some LaTeX defaults for better treatment of figures:
% See p.105 of "TeX Unbound" for suggested values.
% See pp. 199-200 of Lamport's "LaTeX" book for details.
% General parameters, for ALL pages:
\renewcommand{\topfraction}{0.9}    % max fraction of floats at top
\renewcommand{\bottomfraction}{0.8} % max fraction of floats at bottom
% Parameters for TEXT pages (not float pages):
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{2} % 2 may work better
\setcounter{dbltopnumber}{1} % for 2-column pages
\renewcommand{\dbltopfraction}{0.9} % fit big float above 2-col. text
\renewcommand{\textfraction}{0.05}  % allow minimal text w. figs
% Parameters for FLOAT pages (not text pages):
\renewcommand{\floatpagefraction}{0.8}  % require fuller float pages
% N.B.: floatpagefraction MUST be less than topfraction !!
\renewcommand{\dblfloatpagefraction}{0.8}   % require fuller float pages

%
\begin{document}

\title{Does Anybody Really Know What Time It Is?\\
Automating the Extraction of Date Scalars}
\input{anonymous}

\maketitle
\begin{abstract}
Interactive visual analytics is an effective approach for the timely analysis of big data. Users who are already engaged in interactive data analysis prefer to stay in the visualization environment for unavoidable data cleaning and preparation tasks in order to preserve their analytic flow. This has led visualization environments to include simple data preparation functions such as scalar parsing, pattern matching and categorical binning. One common scalar parsing tasks is extracting date and time data from string representations. Several relational database systems include date parsing ``mini-languages'' to cover the wide range of possible formats, but analysis of user data from one visualization system shows that the parsing language syntax is difficult for users to master without disturbing their analytic flow.

In this paper, we present two algorithms for automatically deriving date formats from a column of data with minimal user disruption, one based on minimal entropy representations and another based on natural language processing techniques. Both have accuracies of over 95\% on a large corpus of date columns extracted from an online data repository, and one is also fast enough to produce results within the user's perceptual threshold. Moreover, we were able to avoid prohibitively expensive manual verification by using the algorithms to cross-check each other at scale.
\end{abstract}

%Looks like they don't use them.
%\vidya{Need to figure out the IEEE equivalent for categories and terms}
%D.3.3 [Data Processing]: Data Cleaning, Natural Language Processing.
% A category with the (minimum) three required fields
%\category{D.3.3}{Data Processing}{Data Cleaning, Natural Language Processing}
%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

%\terms{Algorithms, Performance}

\section{Introduction} \label{sec:intro}
\input{intro}

\section{Background} \label{sec:background}
\input{background}

\section{Minimal Descriptive Length} \label{sec:mdl}
\input{mdl}

\section{Natural Language Processing} \label{sec:nlp}
\input{nlp}

\section{Evaluation} \label{sec:experiments}
\input{testing}

\section{Future Work} \label{sec:future}
\input{futurework}

\section{Conclusion} \label{sec:conclusion}
\input{conclusion}

%ACKNOWLEDGMENTS are optional
%\section{Acknowledgments}
%\input{credits}

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{dateparse}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ``.bib'' file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%

%\balancecolumns
% That's all folks!
\end{document}