### 2016-05-13 cikm2016/futurework.tex
minor edits to future work section.

During testing, we found a number of columns that we were not able to process with these techniques. Many columns contained multiple date formats. We would like to recognize this situation and generate predicate-based calculations (\textit{e.g.}, if the string matches format 1 then apply format 1 else ...) to increase the accuracy of the results and thereby make the experience even more seamless for the user. Other columns contain date ranges, which we would like to handle by generating multiple calculations, possibly by combining regular expressions with date parsing.

In this paper, we have considered the parsing of strings, but dates are often formatted as integers (\textit{e.g.}, \functionname{201507016}). 
It is significantly faster to decompose integers into date parts using arithmetic operations (\textit{e.g.}, \texttt{mod} and \texttt{div}) than by using locale-sensitive string parsing functions. 
Timestamp preparation from numeric representations is a related task that we would also like to automate.

In the course of our research, we have also identified a number of date part variants (\textit{e.g.}, ordinal dates, four-letter month abbreviations, alternate meridian markers and postfix quarter syntax) that we would like to commend to the ICU project, along with possible implementations.